Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-11-12 paper/main.tex
work on lp section and analysis soundness

To get a tight bound, we use modern LP solvers that allow constraint
solving and minimization at the same time:  First we consider our initial
constraint set as given in \pref{fig:derivation} and ask the
solver to find a solution that satisfies the constraints and minimizes
the linear expression
$$
1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x} \; .
$$
The penalties given to certain factors are used
to prioritize certain intervals.  For example, a bound with
$[10,x]$ will be prefered to another with $[0,x]$ because $|[10,x]|
\le |[0,x]|$.
%
The LP solver now returns a solution of the constraint set and an
objective value\iffull{, that is, a mapping from variables to floating-point
numbers and the value of the objective function with this
instantiation}{}.  The solver also memorizes the optimization path that
led to the optimal solution.  In this case, the objective value would
be 5000
since the LP solver assigns $q_{0,x} = 0.5$ and $q_{*} = 0$ otherwise.
We now add the constraint
$$
1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
 \le 5000
$$
to our constraint set and ask the solver to optimize the objective
function
$$
q_0 + 11{\cdot}q_{0,10} \; .
$$
This happens in almost no time in practice.
The final solution is $q_{0,x} = 0.5$ and $q_{*} = 0$ otherwise.  Thus
the derived bound is $0.5 |[0,x]|$.

@@

@@

Following the previous work, we define quantitative Hoare triples
as $\Delta;B;R\vdash\htriple Q S {Q'}$ where $B$, $R$, $Q$, and $Q'$ are
maps from program states to $\Qplusz\cup\{\infty\}$ in order to represent
an amount of resources available.  The assertions $B$ and $R$ are
postconditions for the case in which the block $S$ exits by a \code{break}
or \code{return} statement.  Additionally, $R$ depends on an extra value
that is returned by the function.  The meaning of the triple $\htriple Q S
{Q'}$ is as follows: If $S$ is executed with starting state $\state$,
the empty continuation \Kstop{} and at least $Q(\state)$ resource units
available then the evaluation does not run out of resources and there
are at least $Q(\state')$ resources left if the evaluation terminates
in state $\state'$.  The logic rules are similar to the ones in previous
work and are simply extended to account for all the costs found in
our cost-aware semantics.

@@

We translate judgements of the automatic amortized analysis as defined in
\pref{fig:auto} into quantitative Hoare triples.  We define a translation
function $\tr$, such that if a judgement $J$ in the automatic analysis
is derivable, $\tr(J)$ is derivable in the quantitative logic.  By using
$\tr$ to translate derivations of the automatic analysis to derivations
in the quantitative logic we can automatically obtain a certified resource
bound for the analyzed program.

@@

\begin{figure*}
\small
\begin{mathpar}
%
\Rule{L:Update}
{ }
{ \Delta; B; R \vdash
\htriple
  { \lambda \state.\,M_u + M_e(e) + Q\,\state[x \mapsto \sem{e}_\state] }{ x \gets e }{ Q }
}
%
\and
\Rule{L:Assert}
{ }
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state \implies Q + M_a }
  { \code{assert}~e }
  { Q }
}
%
\and
\Rule{L:Tick}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q + M_t(n) }
  { \code{tick}(n) }
  { Q }
}
%
\and
\Rule{L:Loop}
{ \Delta; Q; R \vdash
\htriple
  { I }{ S }{ I + M_l }
}
{ \Delta; B; R \vdash
\htriple
  { I }{ \code{loop}~S }{ Q }
}
\and \Rule{L:Seq}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1 }{ Q' + M_s }
\\
\!\!  \Delta; B; R \vdash
\htriple{ Q' }{ S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1;S_2 }{ Q }
}
%
\and
\Rule{L:Weaken}
{
  \!\! B' \models B \!\! \\
  \!\! \forall v.\, (R'\,v \models R\,v) \!\! \\\\
  \!\! P \models P' \!\! \\
  \!\! \Delta; B'; R' \vdash
  \htriple{ P' }{ S }{ Q' } \!\! \\
  \!\! Q' \models Q \!\!
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\and
\Rule{L:Frame}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q } \\
  x \in \Qplusz
}
{ \!\! \Delta; B + x; R + x \vdash
\htriple{ P + x }{ S }{ Q + x } \!\!
}
%
\end{mathpar}

\caption{Selected rules of the Quantitative Hoare Logic}
\label{fig:logic}
\end{figure*}

>>>>

To get a tight bound, we use modern LP solvers that allow constraint
solving and minimization at the same time:  First we consider our initial
constraint set as given in \pref{fig:derivation} and ask the
solver to find a solution that satisfies the constraints and minimizes
the linear expression \ifshort{$L := 1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x}
+ 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}$.}{
$$
L := 1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x} \; .
$$}
The penalties given to certain factors are used
to prioritize certain intervals.  For example, a bound with
$[10,x]$ will be prefered to another with $[0,x]$ because $|[10,x]|
\le |[0,x]|$.
%
The LP solver now returns a solution of the constraint set and an
objective value\iffull{, that is, a mapping from variables to floating-point
numbers and the value of the objective function with this
instantiation}{}.  The solver also memorizes the optimization path that
led to the optimal solution.  In this case, the objective value would
be 5000
since the LP solver assigns $q_{0,x} = 0.5$ and $q_{*} = 0$ otherwise.
We now add the constraint \ifshort{$L \le 5000$}{
$$
1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
 \le 5000
$$}
to our constraint set and ask the solver to optimize the objective
function \ifshort{$q_0 + 11{\cdot}q_{0,10}$.}{
$$
q_0 + 11{\cdot}q_{0,10} \; .
$$}
This happens in almost no time in practice.
The final solution is $q_{0,x} = 0.5$ and $q_{*} = 0$ otherwise, thus
the bound is $0.5 |[0,x]|$.

@@

\begin{figure*}
\small
\begin{mathpar}
%
\Rule{L:Update}
{ }
{ \Delta; B; R \vdash
\htriple
  { \lambda \state.\,M_u + M_e(e) + Q\,\state[x \mapsto \sem{e}_\state] }{ x \gets e }{ Q }
}
%
\and
\Rule{L:Assert}
{ }
{ \Delta; B; R \vdash
\htriple
  { \code{istrue}~\sem{e}_\state \implies Q + M_a }
  { \code{assert}~e }
  { Q }
}
%
\and
\Rule{L:Tick}
{ }
{ \Delta; B; R \vdash
\htriple
  { Q + M_t(n) }
  { \code{tick}(n) }
  { Q }
}
%
\and
\Rule{L:Loop}
{ \Delta; Q; R \vdash
\htriple
  { I }{ S }{ I + M_l }
}
{ \Delta; B; R \vdash
\htriple
  { I }{ \code{loop}~S }{ Q }
}
\and \Rule{L:Seq}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1 }{ Q' + M_s }
\\
\!\!  \Delta; B; R \vdash
\htriple{ Q' }{ S_2 }{ Q }
}
{ \Delta; B; R \vdash
\htriple{ P }{ S_1;S_2 }{ Q }
}
%
\and
\Rule{L:Weaken}
{
  \!\! B' \models B \!\! \\
  \!\! \forall v.\, (R'\,v \models R\,v) \!\! \\\\
  \!\! P \models P' \!\! \\
  \!\! \Delta; B'; R' \vdash
  \htriple{ P' }{ S }{ Q' } \!\! \\
  \!\! Q' \models Q \!\!
}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q }
}
%
\and
\Rule{L:Frame}
{ \Delta; B; R \vdash
\htriple{ P }{ S }{ Q } \\
  x \in \Qplusz
}
{ \!\! \Delta; B + x; R + x \vdash
\htriple{ P + x }{ S }{ Q + x } \!\!
}
%
\end{mathpar}
\caption{Selected rules of the Quantitative Hoare Logic}
\label{fig:logic}
\end{figure*}

@@

Following the previous work, we define quantitative Hoare triples
as $\Delta;B;R\vdash\htriple Q S {Q'}$ where $B$, $R$, $Q$, and $Q'$ are
maps from program states to $\Qplusz\cup\{\infty\}$ in order to represent
an amount of resources available.  The assertions $B$ and $R$ are
postconditions for the case in which the block $S$ exits by a \code{break}
or \code{return} statement.  Additionally, $R$ depends on an extra value
that is returned by the function.  The meaning of the triple $\htriple Q S
{Q'}$ is as follows: If $S$ is executed with starting state $\state$,
the empty continuation \Kstop{} and at least $Q(\state)$ resource units
available then the evaluation does not run out of resources and there
are at least $Q(\state')$ resources left if the evaluation terminates
in state $\state'$.  The logic rules are similar to the ones in previous
work and are simply extended to account for all the costs found in
our cost-aware semantics.  Some rules are presented in \pref{fig:logic}.

Finally, still following previous work, we define the a continuation based
soundness of triples and prove the soundness of all rules in Coq.  The
full version of this paper \cite{anon_extended}, gives more explanations
on the rules and provides a thorough overview of our soundness proof.

@@

We use the
logic described in the previous paragraph as the target of a translation
function for the automatic derivation system.  This reveals two orthogonal
aspects of our proof that relies, on one side, on amortized reasoning
(as encoded in the quantitative logic), and on the other side, on
combinatorial properties of the linear potential functions that we
are using (as encoded in the automatic analysis).

We define a translation
function $\tr$, such that if a judgement $J$ in the automatic analysis
is derivable, $\tr(J)$ is derivable in the quantitative logic.  By using
$\tr$ to translate derivations of the automatic analysis to derivations
in the quantitative logic we can automatically obtain a certified resource
bound for the analyzed program.

@@
